---
title: "Homework 3"
author: "Megan Wien"
date: "05/28/26"
format: pdf
---

##Set Up

```{r}
#| label: reading-packages
#| message: false
# reading in packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)
```

```{r}
#| label: data-set-up
# creating object for kelp data
kelp <- read_csv("temp-kelp.csv")
# creating object for personal data
commute <- read_csv("envs193ds-personal_data - Sheet1.csv")
```


## Problem 1

### a.

The two tests that we should run to determine the relationship between temp. and giant kelp elongation rate would be a Pearson correlation test and a Spearman rank test. The tests both measure the strength of the relationship between the variables but the Pearson test relies on a normal and linear distribution and the Spearman test does not need normal distribution.

### b. 

```{r}
#| label: problem-1b
# base layer ggplot
ggplot(kelp, mapping = aes(x = temp_c, y = kelp_elong)) + 
  # first layer: scatter plot
  geom_point(color = "blue") + 
  # second layer: line of best fit + ribbon
  geom_smooth(method = "lm", se = TRUE, color = "darkblue") +
  # renaming x and y axis
  labs(x = "Temperature (C)", y = "Kelp Frond Elongation Rate (cm/day)") + 
  # changing the theme of the graph to classic
theme_classic()
```

### c. 
#### Checking assumptions
Linear = yes the data does have a linear relationship. You can see this in problem 1b where the linear graph is the right model for the data. 

Normality = Yes the data is normally distributed, this is shown through the Shapiro test.

Based on these two assumption checks the correct test to run would be the Pearson correlation test. 

```{r}
#| label: checking-normality
shapiro.test(kelp$temp_c) # running normality test for temp
shapiro.test(kelp$kelp_elong) # running normality test for kelp growth rate
```

#### running the test

```{r}
# running the correlation test
cor.test(kelp$temp_c, # setting first variable
         kelp$kelp_elong, # setting 2nd variable
         method = "pearson") # setting the method to be Pearson
```

### d. 

In order to test the strength of the relationship between temperature in Celcius and kelp frond elongation rate in cm/yr I used a Pearson test of correlation (r(30) = -0.687, p = 1.36 x 10^-5^, $\alpha$ = 0.05). This test was ran because the data was normally distributed and showed a linear relationship. This test yielded a value of r = -0.69, which shows a moderate, negative linear correlation between variables. Because of this value and the p value being less than 0.05 we can safely reject the null hypothesis on the grounds of there being a negative correlation between the variables. 















